A Rough Guide to Pre-processing High-Frequency Animal Tracking Data

### 2020-10-08 02_filtering_data.Rmd
make edits to colours etc

```{r}
# define a four colour palette
pal <- RColorBrewer::brewer.pal(4, "Set1")
```

@@

```{r}
# make figure of canonical data with added errors
figure_raw <-
  ggplot()+
  geom_point(data = data_copy,
            aes(x, y),
            col = "grey",
            alpha = 1,
            size = 0.2)+
  geom_path(data = data,
            aes(x, y),
            col = "grey20",
            alpha = 1)+
  ggthemes::theme_map()+
  coord_equal()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "none")+
    theme(plot.background = element_rect(fill = NA))+
  labs(colour = NULL)
```

@@

```{r}
# plot data inside and outside bbox
fig_filter_bounds <-
  ggplot()+
  geom_point(data = data_inside_bbox,
             aes(x, y),
             col = pal[2],
             alpha = 1, size = 0.2)+
  geom_point(data = data_copy[!data_inside_bbox,
                              on = c("x", "y")],
             aes(x, y),
             col = pal[1],
             alpha = 0.5, size = 0.2)+
  geom_path(data = data,
            aes(x, y),
            col = "grey20",
            alpha = 1)+
  geom_hline(yintercept = c(0.5, 1),
             col = "grey",
             lty = 2)+
  ggthemes::theme_map()+
  theme(plot.background = element_rect(fill = NA))+
  coord_equal()

@@

```{r}
# data plot
fig_outlier_remove <-
  ggplot()+
  geom_path(data = data_copy,
            aes(x, y),
            col = "grey",
            lwd = 0.2)+
  geom_point(data = data_copy[500:800, ],
             aes(x, y),
             shape = 2, col = pal[4])+
  geom_point(data = data_copy[!data_copy[500:800, ],
                              on = c("x", "y")],
             aes(x, y, 
                 col = (in_speed >= 0.03 & out_speed >= 0.03),
                 shape = (in_speed >= 0.03 & out_speed >= 0.03)),
             show.legend = F,
             alpha = 0.5)+
  geom_path(data = data,
            aes(x, y),
            col = "grey20",
            alpha = 1)+
  geom_path(data = data,
            aes(x, y),
            col = "grey20",
            alpha = 1)+
  ggthemes::theme_map()+
  scale_color_brewer(palette = "Set1", direction = -1)+
  scale_shape_manual(values = c(1, 19))+
  coord_equal()+
  theme(plot.background = element_rect(fill = NA))
```

@@

```{r}
# attempt to remove reflections
data_no_reflection <- atl_remove_reflections(data_with_reflection,
                          point_angle_cutoff = 10,
                          reflection_speed_cutoff = 0.024)
# get reflections
reflection <- data_with_reflection[!data_no_reflection,
                                         on = c("x", "y")]
reflection <- na.omit(reflection)
```

```{r}
# get plots
fig_reflection <-
  ggplot()+
    geom_path(data = reflection,
             aes(x, y),
             alpha = 1,
             col = "grey")+
  geom_point(data = reflection,
             aes(x, y),
             alpha = 0.5,
             col = pal[1],
             shape = 2)+
  geom_point(data = data_no_reflection,
             aes(x, y),
             alpha = 0.5,
             colour = pal[2],
             shape = 1,
             show.legend = F)+
  geom_path(data = data,
            aes(x, y),
            col = "grey20",
            alpha = 1)+
  ggthemes::theme_map()+
  coord_equal()+
  theme(plot.background = element_rect(fill = NA))

@@

plot_figure

>>>>

```{r}
# define a four colour palette
pal <- RColorBrewer::brewer.pal(3, "PRGn")
```

@@

```{r}
# make figure of canonical data with added errors
figure_raw <-
  ggplot()+
  geom_point(data = data_copy,
            aes(x, y),
            col = "grey",
            alpha = 1,
            size = 0.2)+
  geom_path(data = data,
            aes(x, y),
            col = "grey20",
            alpha = 1)+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "grey99"))+
  coord_equal()+
  labs(colour = NULL)
```

@@

```{r}
# plot data inside and outside bbox
fig_filter_bounds <-
  ggplot()+
  geom_point(data = data_inside_bbox,
             aes(x, y),
             col = pal[3],
             alpha = 1, size = 0.2)+
  geom_point(data = data_copy[!data_inside_bbox,
                              on = c("x", "y")],
             aes(x, y),
             col = "grey",
             alpha = 0.5, size = 0.2)+
  geom_path(data = data,
            aes(x, y),
            col = "grey20",
            alpha = 1)+
  geom_hline(yintercept = c(0.5, 1),
             col = "grey",
             lty = 2)+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "grey99"))+
  theme(plot.background = element_rect(fill = NA))+
  coord_equal(expand = T)

@@

```{r}
# data plot
fig_outlier_remove <-
  ggplot()+
  geom_path(data = data_copy,
            aes(x, y),
            col = "grey",
            lwd = 0.2)+
  geom_point(data = data_copy[500:800, ],
             aes(x, y),
             shape = 2, col = pal[1])+
  geom_point(data = data_copy[!data_copy[500:800, ],
                              on = c("x", "y")],
             aes(x, y, 
                 col = (in_speed >= 0.03 & out_speed >= 0.03),
                 shape = (in_speed >= 0.03 & out_speed >= 0.03)),
             show.legend = F,
             alpha = 0.5)+
  geom_path(data = data,
            aes(x, y),
            col = "grey20",
            alpha = 1)+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "grey99"))+
  scale_color_manual(values = c(pal[3], "black"))+
  scale_shape_manual(values = c(16, 21))+
  coord_equal()+
  theme(plot.background = element_rect(fill = NA))
```

@@

```{r}
# attempt to remove reflections
data_no_reflection <- atl_remove_reflections(data_filtered,
                          point_angle_cutoff = 10,
                          reflection_speed_cutoff = 0.024)
# get reflections
reflection <- data_filtered[!data_no_reflection,
                                         on = c("x", "y")]
reflection <- na.omit(reflection)
```

```{r}
# get plots
fig_reflection <-
  ggplot()+
    geom_path(data = reflection,
             aes(x, y),
             alpha = 1,
             col = "grey",
             lwd = 0.2)+
  geom_point(data = reflection,
             aes(x, y),
             alpha = 0.5,
             col = "grey",
             fill = pal[2],
             shape = 21)+
  geom_point(data = data_no_reflection,
             aes(x, y),
             alpha = 0.5,
             colour = pal[3],
             shape = 16,
             show.legend = F)+
  geom_path(data = data,
            aes(x, y),
            col = "grey20",
            alpha = 1)+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "grey99"))+
  coord_equal()+
  theme(plot.background = element_rect(fill = NA))

@@

### 2020-10-08 03_smoothing_tracks.Rmd
make edits to colours etc

# prepare a palette
pal <- RColorBrewer::brewer.pal(4, "Set1")
```

@@

```{r}
# make a single list of data
data_plot <- append(list(data_errors), list_of_smooths)

# now prep smooth over Y coordinate
fig_smooth_1d <-
  ggplot()+
  geom_point(data = rbindlist(data_plot),
             aes(time, y,
                 col = factor(window_size)),
             alpha = 0.5,
             size = 1,
             show.legend = FALSE,
             shape = 16)+
  coord_cartesian(xlim = c(5000, 7000),
                  ylim = c(0.7, NA),
                  expand = F)+
  scale_colour_brewer(palette = "Spectral")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank())+
  labs(x = "position index", y = "Y coordinate")
```

We prepare a plot of the smoothed data as well as the demonstration of the method relative to gaps.

```{r}
# prepare data to plot
# make list of data to plot
figure_median_smooth <- 
  mapply(function(df, col) {
    ggplot()+
      geom_point(data = df,
                 aes(x, y),
                 size = 0.5,
                 alpha = 0.5,
                 show.legend = FALSE,
                 col = col,
                 shape = 1)+
      geom_path(data = data,
                aes(x, y),
                col = "grey20")+
      coord_cartesian(expand = T,
                  ylim = c(0.6, 0.8)) +
      ggthemes::theme_map()
}, data_plot, c(pal[2], rep(
  RColorBrewer::brewer.pal(3, "Greens")[2],
  4)),
SIMPLIFY = F)

# make basic plot with rectangle
fig_bounds <-
  ggplot()+
  geom_point(data = data_errors,
            aes(x, y),
            size = 0.1, alpha = 0.2, 
            col = "grey")+
  geom_path(data = data,
            aes(x, y),
            size = 0.1)+
  geom_hline(yintercept = c(0.6, 0.8),
             col = "grey20",
             lty = 2, lwd = 0.2)+
  coord_equal(expand = F) +
  ggthemes::theme_map()

# append to first figure
figure_median_smooth[[1]] <-
  figure_median_smooth[[1]]+
  annotation_custom(grob = ggplotGrob(fig_bounds),
                    ymin = 0.6, ymax = 0.8,
                    xmin = 0.5, xmax = 0.7)

# wrap plots
fig_median_smooth <- wrap_plots(append(figure_median_smooth, 
                                       list(fig_smooth_1d)), 
                                       ncol = 3)+
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")") &
  theme(plot.tag = element_text(face = "bold"))

# save figure
ggsave(fig_median_smooth, filename = "figures/fig_median_smooth.png",
       width = 225, height = 225 / 1.77, units = "mm")
```

@@

We shall prepare a figure showing the aggregation steps.

```{r}
# make figure of aggregation
figure_aggregate <- 
  mapply(FUN = function(df) {
  ggplot(data = df)+
      geom_point(data = list_of_smooths[[3]],
                 aes(x, y),
                 col = pal[3],
                 size = 0.3, alpha = 0.5)+
      geom_point(aes(x, y, size = SD),
                 colour = pal[4],
                 alpha = 1,
                 stroke = 1,
                 shape = 4, show.legend = F)+
      geom_path(aes(x, y),
                 colour = pal[4],
                 alpha = 1)+
      geom_path(data = data,
                lwd = 0.2,
                aes(x, y),
                col = "grey20")+
      coord_equal(ylim = c(0.6, 0.8))+
      ggthemes::theme_few()+
      theme(axis.title = element_blank(),
            axis.text = element_blank())
}, list_of_agg[2:3],
SIMPLIFY = FALSE)

figure_aggregate <- wrap_plots(figure_aggregate)+
    plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")") &
    theme(plot.tag = element_text(face = "bold"))

# save figure
ggsave(figure_aggregate, 
       filename = "figures/fig_aggregate_data.png",
       width = 185, height = 90, units = "mm")
```

@@

```{r}
### plot figures
fig_agg_data <- 
  lapply(list_of_agg, function(df) {
    ggplot(df)+
      geom_path(aes(x,y), 
                # col = factor(interval), 
                # group = interval),
                lwd = 0.2)+
      geom_point(aes(x,y, 
                     size = SD,
                     shape = speed < quantile(data$speed, 0.95, na.rm = T),
                     col = speed < quantile(data$speed, 0.95, na.rm = T),
                     group = interval),
                 stroke = ifelse(df$speed < quantile(data$speed, 0.95, na.rm = T),
                                 1, 1),
                 show.legend = F)+
      geom_path(data = data,
                aes(x, y),
                lwd = 0.1)+
      scale_shape_manual(values = c(2, 4))+
      scale_colour_manual(values = pal[c(1, 2)])+
      coord_cartesian(ylim = c(0.6, NA))+
      ggthemes::theme_map()  
  })
```

>>>>

# prepare a palette
pal <- RColorBrewer::brewer.pal(3, "PRGn")
```

@@

```{r}
# bind list after offset
data_plot <- mapply(function(df, offset) {
  df[, x := x + offset]
}, list_of_smooths, seq(0.4, 1.25, length.out = 4),
SIMPLIFY = F)

data_plot <- rbindlist(data_plot)
```

We prepare a plot of the smoothed data.

```{r}
# prepare data to plot
# make list of data to plot
figure_median_smooth <-
  ggplot()+
  geom_point(data = data_errors,
             aes(x, y),
             col = pal[3],
             size = 0.2, alpha = 0.5)+
  geom_path(data = data,
            aes(x, y),
            col = "grey20",
            lwd = 0.5)+          
  geom_path(data = data_plot,
            aes(x, y,
                col = window_size,
                group = window_size),
            show.legend = F,
            lwd = 0.5)+
  coord_equal(expand = F,
              ylim = c(0.6, 0.85),
              xlim = c(NA, 2.3),
              ratio = 1.75)+
  annotate(geom = "text",
           x = c(0.75, seq(1.1, 2, length.out = 4)),
           y = 0.82,
           label = sprintf("(%s)", letters[seq(5)]),
           fontface = "bold")+
  scale_colour_distiller(palette = "BuPu", direction = 1,
                         values = c(-0.5, 1))+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "grey99"))

# # make basic plot with rectangle
# fig_bounds <-
#   ggplot()+
#   geom_point(data = data_errors,
#             aes(x, y),
#             size = 0.1, alpha = 0.2, 
#             col = "grey")+
#   geom_path(data = data,
#             aes(x, y),
#             size = 0.1)+
#   geom_hline(yintercept = c(0.6, 0.8),
#              col = "grey20",
#              lty = 2, lwd = 0.2)+
#   coord_equal(expand = F) +
#   ggthemes::theme_map()

# # append to first figure
# figure_median_smooth[[1]] <-
#   figure_median_smooth[[1]]+
#   annotation_custom(grob = ggplotGrob(fig_bounds),
#                     ymin = 0.6, ymax = 0.8,
#                     xmin = 0.5, xmax = 0.7)

# wrap plots
# fig_median_smooth <- 
  # wrap_plots(figure_median_smooth, 
  #            ncol = 3)+
  # plot_annotation(tag_levels = "a",
  #                 tag_prefix = "(",
  #                 tag_suffix = ")") &
  # theme(plot.tag = element_text(face = "bold"))

# save figure
ggsave(figure_median_smooth, filename = "figures/fig_median_smooth.png",
       width = 170, height = 170 / 3, units = "mm")
```

@@

@@

```{r}
### plot figures
fig_agg_data <- 
  lapply(list_of_agg, function(df) {
    ggplot(df)+
      geom_point(data = data_errors,
                 aes(x, y),
                 size = 0.2,
                 col = pal[3])+
            geom_path(data = data,
                aes(x, y),
                lwd = 0.3)+
      geom_path(aes(x,y), 
                col = pal[1],
                lwd = 0.2)+
      geom_point(aes(x,y, 
                     size = SD,
                     fill = speed,
                     group = interval),
                 shape = 16,
                 show.legend = F)+
      scale_fill_distiller(palette = "Purples", direction = 1,
                             values = c(-0.25, 1))+
        ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "grey99"))+
      coord_cartesian(ylim = c(0.6, NA))
  })
```
